---
title: "Social -- Dargebotene Hand"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

# knitr options

knitr::opts_chunk$set(echo = FALSE)

# load packages
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(lubridate)
library(purrr)
library(janitor)
library(here)
library(readxl)
```

# Import excel data

```{r 'read data'}
dh_path <- Sys.getenv("dh_path")
file_paths <- here(dh_path, list.files(path = dh_path))
dat_in <- map(file_paths, ~read_excel(path = .x, skip = 2))
```

# Data tidying

Prepare clean variables and rename unnamed variables. 

```{r 'clean data'}
clean_data <- function(raw_data){
  data_cleaned <- raw_data %>%
    clean_names() %>%
    rename(gruppe = x1,
           variable = x2) %>% 
    # remove rows with NAs only
    filter_all(any_vars(!is.na(.))) %>% 
    slice(-1) %>% 
    # complete dataframe
    fill(gruppe, .direction = "down") %>% 
     # remove some variables
    filter(!(grepl(pattern = c("Arit|Median|Elt"), x = variable))) %>%
    filter(!(grepl(pattern = c("hnreg|erhil|Sprache"), x = gruppe))) %>% 
    # remove percentage data
    select(-in_prozent)
  return(data_cleaned)
}

dat_clean <- map(dat_in, ~clean_data(raw_data = .x))
```

## Validation 1: check sums

Compare the total number of calls in column `Anzahl` to the sum of calls in variable `Gruppe` for value 'Schicht'.

```{r 'check sums'}
check1 <- function(raw_data, cleaned_data){
  anzahl_anrufe_total <- raw_data %>% 
    slice(1) %>% 
    .$Anzahl
  
  check <- cleaned_data %>%
    filter(gruppe == "Schicht") %>% 
    group_by(gruppe) %>% 
    summarise(
      sum_anzahl = sum(anzahl)
    ) %>% 
    ungroup() %>% 
    mutate(
      check_boolean = anzahl_anrufe_total == sum_anzahl
    )
  
  # Stop script if check is FALSE
  if (check$check_boolean == FALSE) {
    stop("Check 1 of Validation 1 is FALSE. Please evaluate error in code chunk 'check-1'")
    
  } else {
    return(TRUE)
  }
}

# The check function is applied to each file
map2_lgl(.x = dat_in, .y = dat_clean, ~check1(raw_data = .x, cleaned_data = .y)) %>% 
  all()
```

# Add metadata 

Relevant metadata is stored in the file names. This metadata is extracted and added to the cleaned data.

```{r 'extract metadata'}

add_metadata <- function(cleaned_data, path_to_files){
  
  # number of files
  no_files <- length(list.files(path = path_to_files))
  
  # extract file_id, month and year from file names
  dat_filename <- tibble(
    file_id = seq(no_files),
    file_name = list.files(path = path_to_files)) %>% 
    separate(col = file_name, into = c("a", "b", "c", "d"), sep = c(19, 24, 26)) %>% 
    separate(col = d, into = c("e", "f"), sep = ".xlsx") %>% 
    select(-a, -e, -f) %>% 
    mutate(jahr = str_replace(b, pattern = "-", replacement = "")) %>% 
    select(-b) %>% 
    rename("monat" = c) %>% 
    mutate(jahr = str_replace(jahr, "_", "")) %>% 
    split(seq(no_files))
    
  # combine cleaned data and extracted metadata
  dat_combined <- map2(.x = dat_filename, .y = cleaned_data, ~cbind(.x, .y))
  return(dat_combined)
}

dat_final <- add_metadata(dat_clean, dh_path)
```


# Bring monthly data into Covid19-Monitoring format

```{r 'covid19 monitoring format'}
dat_export <- bind_rows(dat_final) %>% 
  rename("value" = anzahl) %>% 
  mutate(monat = as.numeric(monat)) %>% 
  unite(date, c("monat", "jahr"), sep = "-", remove = F) %>% 
  mutate(date = dmy(paste(days_in_month(monat), format(date, format = "%m-%y"), sep = "-"))) %>% 
  select(-file_id, -monat, -jahr) %>% 
  group_by(date, gruppe) %>% 
  # summarize Schicht group
  summarize(variable = if_else(gruppe == "Schicht", "total", variable),
            value = if_else(gruppe == "Schicht", sum(value), value),
            date = date) %>% 
  ungroup() %>% 
  distinct() %>% 
  mutate(gruppe = if_else(gruppe == "Schicht", "Kontaktaufnahmen", gruppe),
         variable = case_when(variable == "Ablehnung, ev. gesperrt" ~ "Schweigeanruf",
                              variable == "Vereinbarung" ~ "Regelmässiger Kontakt",
                              TRUE ~ variable)) %>% 
  unite(variable_long, c("gruppe", "variable"), sep = " ") %>% 
  group_by(date, variable_long) %>% 
  summarize(value = sum(value)) %>% 
  ungroup() %>% 
  complete(date, nesting(variable_long)) %>% 
  mutate(topic = "Soziales",
         location = "ZH",
         unit = "Anzahl",
         source = "Dargebotene Hand",
         update = "monatlich",
         public = "ja",
         description = "https://github.com/statistikZH/covid19monitoring_social_DargeboteneHand/blob/main/README.md") %>% 
  relocate(c(date,value, topic), .before = variable_long) %>% 
  mutate(variable_short = case_when(variable_long == "Kontaktaufnahmen total" ~ "kontaktaufnahmen_total",
                                    variable_long == "Geschlecht Weiblich" ~ "geschlecht_weiblich",
                                    variable_long == "Geschlecht Männlich" ~ "geschlecht_maennlich",
                                    variable_long == "Geschlecht nicht bestimmbar" ~ "geschlecht_unbestimmt",
                                    variable_long == "Alter bis 18 J." ~ "alter_bis_18",
                                    variable_long == "Alter 19 - 40 J." ~ "alter_19_bis_40",
                                    variable_long == "Alter 41 - 65 J." ~ "alter_41_bis_65",
                                    variable_long == "Alter über 65 J." ~ "alter_ue65",
                                    variable_long == "Alter nicht bestimmbar" ~ "alter_unbestimmt",
                                    variable_long == "Kontakthäufigkeit Erster Kontakt" ~ "kontakt_erstmalig",
                                    variable_long == "Kontakthäufigkeit Gelegentlicher, wiederholter Kontakt" ~ "kontakt_gelegentlich",
                                    variable_long == "Kontakthäufigkeit Regelmässiger Kontakt" ~ "kontakt_regelmaessig",
                                    variable_long == "Beratungsinhalt Alltagsbewältigung" ~ "inhalt_alltagsbewaeltigung",
                                    variable_long == "Beratungsinhalt Psychisches Leiden" ~ "inhalt_leiden_psychisch",
                                    variable_long == "Beratungsinhalt Sorge wegen Infektion" ~ "inhalt_sorge_infektion",
                                    variable_long == "Beratungsinhalt Körperliches Leiden" ~ "inhalt_leiden_koerperlich",
                                    variable_long == "Beratungsinhalt Einsamkeit" ~ "inhalt_einsamkeit",
                                    variable_long == "Beratungsinhalt Familie / Erziehung" ~ "inhalt_familie_erziehung",
                                    variable_long == "Beratungsinhalt Beziehung allgemein" ~ "inhalt_beziehung_allgemein",
                                    variable_long == "Beratungsinhalt Paarbeziehung" ~ "inhalt_paarbeziehung",
                                    variable_long == "Beratungsinhalt Arbeit / Ausbildung" ~ "inhalt_arbeit_ausbildung",
                                    variable_long == "Beratungsinhalt Existenzprobleme" ~ "inhalt_existenzprobleme",
                                    variable_long == "Beratungsinhalt Gewalt" ~ "inhalt_gewalt",
                                    variable_long == "Beratungsinhalt Sexualität" ~ "inhalt_sexualitaet",
                                    variable_long == "Beratungsinhalt Spiritualität / Lebenssinn" ~ "inhalt_spiritualitaet_lebenssinn",
                                    variable_long == "Beratungsinhalt Suchtverhalten" ~ "inhalt_suchtverhalten",
                                    variable_long == "Beratungsinhalt Suizidalität" ~ "inhalt_suizidalitaet",
                                    variable_long == "Beratungsinhalt Verlust / Trauer / Tod" ~ "inhalt_verlust_trauer_tod",
                                    variable_long == "Beratungsinhalt Verschiedenes" ~ "inhalt_verschiedenes",
                                    variable_long == "Beratungsinhalt Medienerziehung/Swisscom" ~ "inhalt_medienerziehung_swisscom",
                                    variable_long == "Zusätzliche Beanspruchung Verschobenes Gespräch" ~ "zus_beanspruchung_verschobenes_gespraech",
                                    variable_long == "Zusätzliche Beanspruchung Schweigeanruf" ~ "zus_beanspruchung_schweigeanruf",
                                    variable_long == "Zusätzliche Beanspruchung Fehlanruf" ~ "zus_beanspruchung_fehlanruf",
                                    variable_long == "Zusätzliche Beanspruchung Juxanruf" ~ "zus_beanspruchung_juxanruf",
                                    variable_long == "Zusätzliche Beanspruchung aufgelegt" ~ "zus_beanspruchung_aufgelegt",
                                    variable_long == "Geschlecht Divers" ~ "geschlecht_divers",
                                    variable_long == "Zusätzliche Beanspruchung Chat ohne Inhalt" ~ "zus_beanspruchung_chat_ohne_inhalt",
                                    TRUE ~ "no_case"
                                    )) %>% 
  relocate(variable_short, .after = variable_long) 
```

## Validation 2: check variable_short

```{r 'checkk variable_short'}
# check whether there are any "no_case" variables
if(any(dat_export$variable_short == "no_case")){
  stop("Some variables do not have a 'variable_short' definition!")
}
```


# Data Export

Write CSV for monthly data.

```{r 'write data'}
write_csv(dat_export,  here("Social_DargeboteneHand.csv"))
```
